---
title: "PRS_EA"
output: html_document
author: Sebastian Marten & Amaury Pelzer
---

```{r, include=FALSE}
if (suppressWarnings(!require(tidyverse))) {install.packages("tidyverse"); library(tidyverse)}
if (suppressWarnings(!require(ggplot2))) {install.packages("ggplot2"); library(ggplot2)}
if (suppressWarnings(!require(data.table))) {install.packages("data.table"); library(data.table)}
if (suppressWarnings(!require(formattable))) {install.packages("formattable"); library(formattable)}
```
# Test of Time dependency
Variables that are not required for the analysis: "SITE","COLPROT","ORIGPROT","FLDSTRENG","FSVERSION","IMAGEUID", "Month_bl","Month","M","update_stamp".

```{r}
knitr::opts_chunk$set(echo = TRUE)

dat <- read.csv("data/ADNIMERGE.csv")

time_test <- function(data, idvar, tvar) {
  x <- data[, c(idvar, tvar)] %>% arrange(tvar)
  time_dependent <- FALSE
  unique_participants <- unique(x[, 1])
  for (participant in unique_participants) {
    participant_data <- x[x[, 1] == participant, 2]
    if (length(unique(participant_data)) > 1) {
      time_dependent <- TRUE
      break
    }
  }
  if (time_dependent) {
    return(data.frame(variable = tvar, status = "Time Dependent"))
  } else {
    return(data.frame(variable = tvar, status = "Time Independent"))
  }
}

results <- data.frame(variable = character(), status = character(), stringsAsFactors = FALSE)

for (col in 1:length(colnames(dat))) {
  result <- time_test(dat, "RID", colnames(dat)[col])
  results <- rbind(results, result)
}


excluded_vars <- c("SITE","COLPROT","ORIGPROT","FLDSTRENG",
                   "FSVERSION","IMAGEUID", "Month_bl","Month","M",
                   "update_stamp", "FSVERSION_bl", "FLDSTRENG_bl", 
                   "EXAMDATE_bl", "EXAMDATE", "Years_bl")

ivars <- subset(results, status=="Time Independent") %>%
  subset(!(variable %in% excluded_vars)) %>% subset(select=variable)

nivars <- subset(results, status=="Time Dependent") %>% 
  subset(!(variable %in% excluded_vars)) %>% subset(select=variable)

rm(results, result, col)
```

# Data Format: Long to Wide
Here are merge time dependent and independent variables into the long_dat data
frame. Since the tidyverse package has some problems with ordering the VISCODE 
data properly due to the occurrence of both strings and integers we rearrange 
the variable to an only integer format. This will likely also make the 
integration of "..._bl" variables into the multivariate data format more easy.
Thus, we mutate the VISCODE data into integer values, where 0 equal baseline 
measurements, 1 equals month 2 and so on.

```{r}
long_dat <- dat[, c(ivars[,1], nivars[,1])] %>%
  mutate(VISCODE = match(VISCODE, c("bl", "m03", "m06", "m12", "m18", "m24", 
                                    "m30","m36", "m42", "m48", "m54", "m60", 
                                    "m66", "m72","m78", "m84", "m90", "m96", 
                                    "m102", "m108","m114", "m120", "m126", 
                                    "m132", "m144", "m156"))-1) %>%
  relocate(RID, PTID, VISCODE) %>%
  arrange(RID, VISCODE)
```

# _BL variables already correct?
Here I test whether the baseline values maybe have already been integrated 
because for the first few participants that looked correct. However, there 
seem to be discrepancies in some of the values.

```{r, echo = FALSE}
baseline_vars <- long_dat %>%
  select(c("VISCODE","PTID",ends_with("_bl"),ends_with("_BL"))) %>%
  filter(VISCODE == 0) %>%
  subset(select = -c(DX_bl, VISCODE)) %>%
  rename(LDELTOTAL_bl = LDELTOTAL_BL)

baseline_commplementary_vars <- long_dat %>% 
  
  subset(select = c(VISCODE, DX, CDRSB, ADAS11, ADAS13, ADASQ4,MMSE,
  RAVLT_immediate,RAVLT_learning,RAVLT_forgetting,RAVLT_perc_forgetting,
  LDELTOTAL,DIGITSCOR,TRABSCOR,FAQ,mPACCdigit,mPACCtrailsB,Ventricles,
  Hippocampus,WholeBrain,Entorhinal,Fusiform,MidTemp,ICV,MOCA,EcogPtMem,
  EcogPtLang,EcogPtVisspat,EcogPtPlan,EcogPtOrgan,EcogPtDivatt,EcogPtTotal,
  EcogSPMem,EcogSPLang,EcogSPVisspat,EcogSPPlan,EcogSPOrgan,EcogSPDivatt,
  EcogSPTotal,ABETA,TAU,PTAU,FDG,PIB,AV45)) %>%
  
  filter(VISCODE == 0) %>% 
  subset(select = -c(DX, VISCODE))

length(colnames(baseline_vars)) == length(colnames(baseline_commplementary_vars))
all.equal(baseline_vars, baseline_commplementary_vars)
```

# Merge _bl/_BL variables with corresponding time dependent variable

```{r}
baseline_commplementary_col_names <- colnames(baseline_commplementary_vars)

long_dat <- long_dat %>%
  left_join(baseline_vars, by = "PTID", suffix = c("", "_bl")) %>%
  mutate(across(all_of(baseline_commplementary_col_names), ~ coalesce(., get(paste0(cur_column(), "_bl"))))) %>%
  select(-ends_with("_bl"), -ends_with("_BL")) %>%
  mutate(
    RID = as.factor(RID),
    PTID = as.character(PTID),
    VISCODE = as.factor(VISCODE),
    AGE = as.numeric(AGE),
    PTGENDER = as.factor(PTGENDER),
    PTEDUCAT = as.integer(PTEDUCAT),
    PTETHCAT = as.factor(PTETHCAT),
    PTRACCAT = as.factor(PTRACCAT),
    PTMARRY = as.factor(PTMARRY),
    APOE4 = as.integer(APOE4),
    FDG = as.numeric(FDG),
    PIB = as.numeric(PIB),
    AV45 = as.numeric(AV45),
    ABETA = as.numeric(ABETA),
    TAU = as.numeric(TAU),
    PTAU = as.numeric(PTAU),
    CDRSB = as.numeric(CDRSB),
    ADAS11 = as.numeric(ADAS11),
    ADAS13 = as.numeric(ADAS13),
    ADASQ4 = as.integer(ADASQ4),
    MMSE = as.integer(MMSE),
    RAVLT_immediate = as.integer(RAVLT_immediate),
    RAVLT_learning = as.integer(RAVLT_learning),
    RAVLT_forgetting = as.integer(RAVLT_forgetting),
    RAVLT_perc_forgetting = as.numeric(RAVLT_perc_forgetting),
    LDELTOTAL = as.integer(LDELTOTAL),
    DIGITSCOR = as.integer(DIGITSCOR),
    TRABSCOR = as.integer(TRABSCOR),
    FAQ = as.integer(FAQ),
    MOCA = as.integer(MOCA),
    EcogPtMem = as.numeric(EcogPtMem),
    EcogPtLang = as.numeric(EcogPtLang),
    EcogPtVisspat = as.numeric(EcogPtVisspat),
    EcogPtPlan = as.numeric(EcogPtPlan),
    EcogPtOrgan = as.numeric(EcogPtOrgan),
    EcogPtDivatt = as.numeric(EcogPtDivatt),
    EcogPtTotal = as.numeric(EcogPtTotal),
    EcogSPMem = as.numeric(EcogSPMem),
    EcogSPLang = as.numeric(EcogSPLang),
    EcogSPVisspat = as.numeric(EcogSPVisspat),
    EcogSPPlan = as.numeric(EcogSPPlan),
    EcogSPOrgan = as.numeric(EcogSPOrgan),
    EcogSPDivatt = as.numeric(EcogSPDivatt),
    EcogSPTotal = as.numeric(EcogSPTotal),
    Ventricles = as.integer(Ventricles),
    Hippocampus = as.integer(Hippocampus),
    WholeBrain = as.integer(WholeBrain),
    Entorhinal = as.integer(Entorhinal),
    Fusiform = as.integer(Fusiform),
    MidTemp = as.integer(MidTemp),
    ICV = as.integer(ICV),
    DX = as.factor(DX),
    mPACCdigit = as.numeric(mPACCdigit),
    mPACCtrailsB = as.numeric(mPACCtrailsB)
    )
```

# Attrition Analysis
To see how many measurements we have at baseline

```{r, fig.align="center"}
frequency_table <- long_dat %>%
  group_by(VISCODE) %>%
  summarise(NumParticipants = n_distinct(RID))

barplot(frequency_table$NumParticipants~frequency_table$VISCODE, xlab="Time", ylab="Number of Patients Measured",
     main="Attrition Plot")
```
#Transform Long to Wide Data Format
Next we transform the data using the pivot_wider function, which requires three 
inputs. First, a variable to create new names from, which is in our case the 
VISCODE variable of length 26. 

```{r}
df <- pivot_wider(
  long_dat,
  id_cols = as.character(ivars$variable),
  names_from = "VISCODE",
  values_from = as.character(nivars$variable),
  values_fn = list(n = n_distinct),
  names_sep = "_"
)
```

# Code Island

For code that was useful on the way but not any more, though too cool to delete.
Thus, I allowed these code snippets to enjoy retirement on Code Island.

## Index Column based on label
```{r}
grep("RID", colnames(dat))
```

## How many participants do we have?
```{r}
length(unique(dat$RID))
```
