---
title: "PGS Educational Attainment - Preprocessing"
author: Sebastian Marten & Amaury Pelzer
output: rmdformats::robobook
---

```{r, include=FALSE}
if (suppressWarnings(!require(tidyverse))) {install.packages("tidyverse"); library(tidyverse)}
if (suppressWarnings(!require(ggplot2))) {install.packages("ggplot2"); library(ggplot2)}
if (suppressWarnings(!require(data.table))) {install.packages("data.table"); library(data.table)}
if (suppressWarnings(!require(formattable))) {install.packages("formattable"); library(formattable)}
if (suppressWarnings(!require(survival))) {install.packages("survival"); library(survival)}
if (suppressWarnings(!require(ggsurvfit))) {install.packages("ggsurvfit"); library(ggsurvfit)}
if (suppressWarnings(!require(performance))) {install.packages("performance"); library(performance)}
if (suppressWarnings(!require(cowplot))) {install.packages("cowplot"); library(cowplot)}
if (suppressWarnings(!require(NormPsy))) {install.packages("NormPsy"); library(NormPsy)}
if (suppressWarnings(!require(gridExtra))) {install.packages("gridExtra"); library(gridExtra)}
if (suppressWarnings(!require(rmdformats))) {install.packages("rmdformats"); library(rmdformats)}
if (suppressWarnings(!require(nlme))) {install.packages("nlme"); library(nlme)}

```

# Data Preprocessing

1.  First, I wanted to investigate which variables are time dependent and also exclude some that were clearly unnecessary (i.e., "SITE","COLPROT","ORIGPROT", "FLDSTRENG","FSVERSION","IMAGEUID", "Month_bl","Month","M","update_stamp").

```{r, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

dat <- read.csv("data/ADNIMERGE.csv")

time_test <- function(data, idvar, tvar) {
  x <- data[, c(idvar, tvar)] %>% arrange(tvar)
  time_dependent <- FALSE
  unique_participants <- unique(x[, 1])
  for (participant in unique_participants) {
    participant_data <- x[x[, 1] == participant, 2]
    if (length(unique(participant_data)) > 1) {
      time_dependent <- TRUE
      break
    }
  }
  if (time_dependent) {
    return(data.frame(variable = tvar, status = "Time Dependent"))
  } else {
    return(data.frame(variable = tvar, status = "Time Independent"))
  }
}

results <- data.frame(variable = character(), status = character(), stringsAsFactors = FALSE)

for (col in 1:length(colnames(dat))) {
  result <- time_test(dat, "RID", colnames(dat)[col])
  results <- rbind(results, result)
}


excluded_vars <- c("SITE","COLPROT","ORIGPROT","FLDSTRENG",
                   "FSVERSION","IMAGEUID", "Month_bl","Month","M",
                   "update_stamp", "FSVERSION_bl", "FLDSTRENG_bl", 
                   "EXAMDATE_bl", "EXAMDATE", "Years_bl")

ivars <- subset(results, status=="Time Independent") %>%
  subset(!(variable %in% excluded_vars)) %>% subset(select=variable)

nivars <- subset(results, status=="Time Dependent") %>% 
  subset(!(variable %in% excluded_vars)) %>% subset(select=variable)

rm(results, result, col)
```

2.  Merge time dependent and independent variables into the long_dat data frame. Also, I recoded the time points in the VISCODE variable into integers.

```{r}
long_dat <- dat[, c(ivars[,1], nivars[,1])] %>%
  mutate(VISCODE = match(VISCODE, c("bl", "m03", "m06", "m12", "m18", "m24", 
                                    "m30","m36", "m42", "m48", "m54", "m60", 
                                    "m66", "m72","m78", "m84", "m90", "m96", 
                                    "m102", "m108","m114", "m120", "m126", 
                                    "m132", "m144", "m156"))-1) %>%
  relocate(RID, PTID, VISCODE) %>%
  arrange(RID, VISCODE)
```

3.  In the original data frame there were quite some \_bl or \_BL variables. Thus, I wanted to check whether these columns had already been integrated or not at each corresponding time point for each participant. Surprise, the test was negative.

```{r, include= FALSE}
baseline_vars <- long_dat %>%
  select(c("VISCODE","PTID",ends_with("_bl"),ends_with("_BL"))) %>%
  filter(VISCODE == 0) %>%
  subset(select = -c(DX_bl, VISCODE)) %>%
  rename(LDELTOTAL_bl = LDELTOTAL_BL)

baseline_commplementary_vars <- long_dat %>% 
  
  subset(select = c(VISCODE, DX, CDRSB, ADAS11, ADAS13, ADASQ4,MMSE,
  RAVLT_immediate,RAVLT_learning,RAVLT_forgetting,RAVLT_perc_forgetting,
  LDELTOTAL,DIGITSCOR,TRABSCOR,FAQ,mPACCdigit,mPACCtrailsB,Ventricles,
  Hippocampus,WholeBrain,Entorhinal,Fusiform,MidTemp,ICV,MOCA,EcogPtMem,
  EcogPtLang,EcogPtVisspat,EcogPtPlan,EcogPtOrgan,EcogPtDivatt,EcogPtTotal,
  EcogSPMem,EcogSPLang,EcogSPVisspat,EcogSPPlan,EcogSPOrgan,EcogSPDivatt,
  EcogSPTotal,ABETA,TAU,PTAU,FDG,PIB,AV45)) %>%
  
  filter(VISCODE == 0) %>% 
  subset(select = -c(DX, VISCODE))

length(colnames(baseline_vars)) == length(colnames(baseline_commplementary_vars))
all.equal(baseline_vars, baseline_commplementary_vars)
```

4.  Therefore, I continued with merging the \_bl/\_BL variables with the corresponding time dependent variable for each participant. Additionally, I specified the data type of each variable individually for optimal control and oversight over the data structure.

```{r, include= FALSE}
baseline_commplementary_col_names <- colnames(baseline_commplementary_vars)

long_dat <- long_dat %>%
  left_join(baseline_vars, by = "PTID", suffix = c("", "_bl")) %>%
  mutate(across(all_of(baseline_commplementary_col_names), ~ coalesce(., get(paste0(cur_column(), "_bl"))))) %>%
  select(-ends_with("_bl"), -ends_with("_BL")) %>%
  mutate(
    RID = as.factor(RID),
    PTID = as.character(PTID),
    VISCODE = as.numeric(VISCODE),
    AGE = as.numeric(AGE),
    PTGENDER = as.factor(PTGENDER),
    PTEDUCAT = as.integer(PTEDUCAT),
    PTETHCAT = as.factor(PTETHCAT),
    PTRACCAT = as.factor(PTRACCAT),
    PTMARRY = as.factor(PTMARRY),
    APOE4 = as.integer(APOE4),
    FDG = as.numeric(FDG),
    PIB = as.numeric(PIB),
    AV45 = as.numeric(AV45),
    ABETA = as.numeric(ABETA),
    TAU = as.numeric(TAU),
    PTAU = as.numeric(PTAU),
    CDRSB = as.numeric(CDRSB),
    ADAS11 = as.numeric(ADAS11),
    ADAS13 = as.numeric(ADAS13),
    ADASQ4 = as.integer(ADASQ4),
    MMSE = as.integer(MMSE),
    RAVLT_immediate = as.integer(RAVLT_immediate),
    RAVLT_learning = as.integer(RAVLT_learning),
    RAVLT_forgetting = as.integer(RAVLT_forgetting),
    RAVLT_perc_forgetting = as.numeric(RAVLT_perc_forgetting),
    LDELTOTAL = as.integer(LDELTOTAL),
    DIGITSCOR = as.integer(DIGITSCOR),
    TRABSCOR = as.integer(TRABSCOR),
    FAQ = as.integer(FAQ),
    MOCA = as.integer(MOCA),
    EcogPtMem = as.numeric(EcogPtMem),
    EcogPtLang = as.numeric(EcogPtLang),
    EcogPtVisspat = as.numeric(EcogPtVisspat),
    EcogPtPlan = as.numeric(EcogPtPlan),
    EcogPtOrgan = as.numeric(EcogPtOrgan),
    EcogPtDivatt = as.numeric(EcogPtDivatt),
    EcogPtTotal = as.numeric(EcogPtTotal),
    EcogSPMem = as.numeric(EcogSPMem),
    EcogSPLang = as.numeric(EcogSPLang),
    EcogSPVisspat = as.numeric(EcogSPVisspat),
    EcogSPPlan = as.numeric(EcogSPPlan),
    EcogSPOrgan = as.numeric(EcogSPOrgan),
    EcogSPDivatt = as.numeric(EcogSPDivatt),
    EcogSPTotal = as.numeric(EcogSPTotal),
    Ventricles = as.integer(Ventricles),
    Hippocampus = as.integer(Hippocampus),
    WholeBrain = as.integer(WholeBrain),
    Entorhinal = as.integer(Entorhinal),
    Fusiform = as.integer(Fusiform),
    MidTemp = as.integer(MidTemp),
    ICV = as.integer(ICV),
    DX = as.factor(DX),
    mPACCdigit = as.numeric(mPACCdigit),
    mPACCtrailsB = as.numeric(mPACCtrailsB)
    )
```

5.  Transform Long to Wide Data Format

```{r, echo=FALSE}
ivars <- ivars %>% filter(!str_detect(variable, "_bl|_BL") & variable != "DX_bl") %>%
  pull(variable) %>%
  as.character()

nivars <- nivars %>% 
  filter(! str_detect(variable, "VISCODE")) %>%
  pull(variable) %>%
  as.character()

wide_dat <- pivot_wider(
  long_dat,
  id_cols = all_of(ivars),
  names_from = "VISCODE",
  values_from = all_of(nivars),
  values_fn = list(n = n_distinct),
  names_sep = "_"
  )

head(wide_dat)
```

# Age Distribution in Data Frame

```{r, echo=FALSE, warning=FALSE}

# Create histogram of general age range in data set

ggplot(long_dat %>% distinct(RID, .keep_all = TRUE), aes(x = AGE)) +
  geom_histogram(binwidth = 1, fill = "#bbceed", color = "black") +
  geom_vline(xintercept = c(60, 70), linetype = "dashed", color = "#940914", size=1) +  # vertical lines age range
  geom_vline(xintercept = c(65, 75), linetype = "dashed", color = "#940914", size=1) +  # vertical lines age range
  labs(title = "Histogram Age Distribution",
       x = "Age",
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),  # Bold and centred title
        axis.line = element_line(size = 0.5))  # Thicker axis lines

# create data frames of two age groups

age_60to70_df <- long_dat %>% # Data frame filtered for age
  filter(AGE >= 60 & AGE <= 70)

age_65to75_df <- long_dat %>% # Data frame filtered for age
  filter(AGE >= 65 & AGE <= 75)

age_60to75_df <- long_dat %>% # Data frame filtered for age
  filter(AGE >= 60 & AGE <= 75)

# AGE Groups (1,2,3 for 60-65, 66-70, 71-75)

long_dat$Age_Group <- ifelse(
  long_dat$AGE >= 60.0 & long_dat$AGE <= 65.0, 1,
  ifelse(
    long_dat$AGE >= 66.0 & long_dat$AGE <= 70.0, 2,
    ifelse(
      long_dat$AGE >= 71.0 & long_dat$AGE <= 75.0, 3,
      NA_integer_
    )
  )
)

# Append to Wide data frame too 
wide_dat <- wide_dat %>%
  left_join(
    long_dat %>% 
      group_by(RID) %>%
      filter(row_number() == 1) %>%  # Keep only the first row for each unique RID
      select(RID, Age_Group),
    by = "RID"
  )
```

# Attrition Analysis

Based on the number of participants measured at any time point I made a frequency plot to get a first idea of the sampling frequency.

```{r, echo=FALSE, fig.align="center"}
frequency_table <- long_dat %>%
  group_by(VISCODE) %>%
  summarise(NumParticipants = n_distinct(RID))

barplot(frequency_table$NumParticipants~frequency_table$VISCODE, xlab="Time", ylab="Number of Patients Measured",
     main="Attrition Plot")
```

# Domains

## Demographics

```{r, echo=FALSE}
df_demographics <- long_dat %>%
  select(RID, PTID, VISCODE, AGE, PTGENDER, PTEDUCAT, PTETHCAT, PTRACCAT, PTMARRY)

freq_dem <- df_demographics %>%
  group_by(VISCODE) %>%
  summarise(across(everything(), ~ sum(!is.na(.))))

freq_dem_long <- pivot_longer(freq_dem, cols = -VISCODE, names_to = "Variable", values_to = "Frequency")

ggplot(freq_dem_long, aes(x = VISCODE, y = Frequency, color = Variable, group = Variable)) +
  geom_line() +
  labs(x = "Timepoint (VISCODE)", y = "Number of Measurements", title = "Number of Measurements per Time Point") +
  theme(legend.position = "top")
```

## Cognitive Tests

```{r, echo=FALSE}
df_cognitive_tests <- long_dat %>%
  select(RID, PTID, VISCODE, CDRSB, ADAS11, ADAS13, ADASQ4, MMSE,
         RAVLT_immediate, RAVLT_learning, RAVLT_forgetting, RAVLT_perc_forgetting,
         LDELTOTAL, DIGITSCOR, TRABSCOR, FAQ, MOCA,
         EcogPtMem, EcogPtLang, EcogPtVisspat, EcogPtPlan, EcogPtOrgan, EcogPtDivatt, EcogPtTotal,
         EcogSPMem, EcogSPLang, EcogSPVisspat, EcogSPPlan, EcogSPOrgan, EcogSPDivatt, EcogSPTotal)

freq_cog <- df_cognitive_tests %>%
  group_by(VISCODE) %>%
  summarise(across(everything(), ~ sum(!is.na(.))))

freq_cog_long <- pivot_longer(freq_cog, cols = -VISCODE, names_to = "Variable", values_to = "Frequency")

ggplot(freq_cog_long, aes(x = VISCODE, y = Frequency, color = Variable, group = Variable)) +
  geom_line() +
  labs(x = "Timepoint (VISCODE)", y = "Number of Measurements", title = "Number of Measurements per Time Point") +
  theme(legend.position = "top")
```

## Biomedical Imaging

```{r,echo=FALSE}
df_biomedical_imaging <- long_dat %>%
  select(RID, PTID, VISCODE, Ventricles, Hippocampus, WholeBrain, Entorhinal, Fusiform, MidTemp, ICV)

freq_imag <- df_biomedical_imaging %>%
  group_by(VISCODE) %>%
  summarise(across(everything(), ~ sum(!is.na(.))))

freq_imag_long <- pivot_longer(freq_imag, cols = -VISCODE, names_to = "Variable", values_to = "Frequency")

ggplot(freq_imag_long, aes(x = VISCODE, y = Frequency, color = Variable, group = Variable)) +
  geom_line() +
  labs(x = "Timepoint (VISCODE)", y = "Number of Measurements", title = "Number of Measurements per Time Point") +
  theme(legend.position = "top")
```

## Biomarkers

```{r, echo=FALSE}
df_biomarkers <- long_dat %>%
  select(RID, PTID, VISCODE, APOE4, ABETA, TAU, PTAU, FDG, PIB, AV45)

freq_mark <- df_biomarkers %>%
  group_by(VISCODE) %>%
  summarise(across(everything(), ~ sum(!is.na(.))))

freq_mark_long <- pivot_longer(freq_mark, cols = -VISCODE, names_to = "Variable", values_to = "Frequency")

ggplot(freq_mark_long, aes(x = VISCODE, y = Frequency, color = Variable, group = Variable)) +
  geom_line() +
  labs(x = "Timepoint (VISCODE)", y = "Number of Measurements", title = "Number of Measurements per Time Point") +
  theme(legend.position = "top")
```

Based on these findings it appears that time point 9 is a cut-off where the number of measurements drop quite strongly. Time point 9 corresponds to month 42 (i.e., 3.5 years) of the follow-up.

```{r, echo=FALSE}
rm(list= setdiff(ls(), c("long_dat", "wide_dat")))
```

# Polygenic Score for Educational attainment

```{r, echo=FALSE, message=FALSE}
# Read in the PGS send to us by Rick
pgs <- readr::read_tsv("data/ADNI_PRS_results_EA_EA22.tsv")

# Merge the EA data with the corresponding participant in each data frame
long_dat <- merge(long_dat, pgs,by.x="PTID",by.y="SampleID") 
wide_dat <- merge(wide_dat, pgs,by.x="PTID",by.y="SampleID")
```

The merge(by.x, by.y) function creates a new data frame that only keeps those rows for which there is a matching key (in our case PTID). Therefore, we do have genetic data from 2 additional individuals for which we do not have any other measurements. The final data frame for which testing data and genetic data is available is thus, 1408 (N).

## Plot PGS EA vs. Actual EA

Based on this plot, we can see a positive relationship between the polygenic score for education attainment and actual years of education. This means that with a higher PGS score comes higher genetic capacity for educational attainment.

We ran Pearson's correlation which resulted in r = 0.286 (p-value \< 2.2e-16)

```{r, echo=FALSE, message=FALSE, results='hide'}
# Plot PGS EA against Actual EA
ggplot(long_dat, aes(x = PTEDUCAT, y = EA22)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm") +  # Add linear regression line
  labs(x = "Years of Education", y = "EA22") +  # Set axis labels
  ggtitle("EA22 vs Years of Education")  # Set plot title

cor(long_dat$PTEDUCAT, long_dat$EA22, method="pearson")
cor.test(long_dat$PTEDUCAT, long_dat$EA22, method="pearson")
```

## Check linear regression assumptions

```{r, echo=FALSE}
lm(PTEDUCAT~PTGENDER, data=wide_dat)
boxplot(PTEDUCAT~PTGENDER, data=wide_dat)

lm(EA22~PTGENDER, data=wide_dat)
boxplot(EA22~PTGENDER, data=wide_dat)

# compare n gender
table(wide_dat$PTGENDER)

# Linear regression model
model <- lm(PTEDUCAT~EA22+AGE+PTGENDER + AGE*PTGENDER,data=wide_dat)

# Check model assumptions
check_model(model)
```

## Create Residuals

To get the residual we regressed the polygenic risk score for educational attainment against actual EA including the variables SEX & AGE as covariates. The results are depicted in the density plot.

```{r, echo=FALSE}
# Create the residuals given two covariates AGE and Sex
results <- lm(PTEDUCAT~EA22+AGE+PTGENDER+AGE*PTGENDER,data=wide_dat)

# Standardize the residuals and append to data frame
wide_dat$res <- rstandard(results)

# Create density plot of the residuals
ggplot(wide_dat, aes(x = res)) +
  geom_density(fill = "orange", alpha = 0.5) +
  labs(x = "Residual Score", y = "Density") +
  ggtitle("Distribution of Residuals")
```

## How to interpret the Residuals?

It is important to correctly interpret the residual scores. The correct way to interpret them is, that a high residual score means that the individual has over-performed relative to his or her genetic capacity. See for example in this table for a short proof:

```{r, echo=FALSE}
residuals <- results$residuals # get residuals
predicted_scores <- predict(results) # get predicted scores for each individual

combined_df <- data.frame(Actual = wide_dat$PTEDUCAT, Predicted = as.numeric(predicted_scores), Residuals = as.numeric(residuals)) # Combine residuals, predicted scores, and actual scores into one data frame

head(combined_df %>% filter(!duplicated(Actual))) # present data frame as example
```

```{r, echo=FALSE}
# Create a residual plot
ggplot() +
  geom_qq(aes(sample = residuals)) +
  geom_qq_line() +
  labs(title = "Normal Q-Q Plot of Residuals") +
  theme_minimal()
```

```{r, eval=FALSE}
# Create a folder called "output" if it doesn't exist
if (!file.exists("output")) {
  dir.create("output")
}

# Save wide_dat as a CSV file in the "output" folder
write.csv(wide_dat, file = "output/wide_dat.csv", row.names = FALSE)

# Save long_dat as a CSV file in the "output" folder
write.csv(long_dat, file = "output/long_dat.csv", row.names = FALSE)
```

## Print session info:

```{r print_session_info}
##Print session info:
sessionInfo()
```
